```{r, message=FALSE, echo=FALSE}
library(ggplot2)
```
---
title: "ToothGrowth Exploratory Analysis"
author: "Axel Garcia"
date: "21 de junio de 2015"
output: html_document
---
## Overview
This document discusses the analysis done on the ToothGrowth data provided in R.
That data set was obtained from C. I. Bliss (1952) The Statistics of Bioassay. 
Academic Press (according with the R Documentation).

That data set is mean to describe the effect of vitamin C on tooth growth in Guinea
pigs. More information about this data can be found in the R Documentation. The
Following link points to it: [The Effect of Vitamin C on Tooth Growth in Guinea Pigs](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/ToothGrowth.html)

## Data Information
Format

A data frame with 60 observations on 3 variables.

[,1]    len	numeric	Tooth length

[,2]	supp	factor	Supplement type (VC or OJ).

[,3]	dose	numeric	Dose in milligrams.

```{r Summary}
data(ToothGrowth)

str(ToothGrowth)
```

Supp is a factor variable whereas the other two are numeric variables; furthermore, 
those numeric variables are continuous. 

Wondering about the data, it maybe that the dose has something to do with the teeth 
length. The following plot shows how the relation ship between those two variables 
looks like:
```{r,fig.height=3}
ggplot( ToothGrowth, aes(y=len, x=dose) ) + 
    geom_boxplot() + facet_grid( .~supp) +
    ggtitle( "Dose effect on Length by Supplement Type" )
```
The mean of both variables is contained by the 2nd and 3rd quartiles. That means
that both data overlaps. Given that evidence, the dose does not seem to have a
significaive effect on the teeth length.

In the other hand, the supplement type may have an impact on the teeth lenght. Lets see how the 
relationship beteen those two variables look like.
```{r,fig.height=3}
ggplot( ToothGrowth, aes(y=len, x=supp) ) + 
    geom_boxplot() + facet_grid( .~dose) +
    ggtitle( "Supplement Type effect on Length by Dose" )
```

When the dose is 0.5 the chart apparently shows that the means of OJ and VC 
supplement types are not contained in each other. However, there is still an
overlap between the 1st quartile of OJ and the 3rd quartile of VC. Lets see
what additional information the hypothesis test provides for this case.

When dose is 2, the plot shows that OJ data is completely with in the VC
distribution. In this case it does not seem to be any difference between them.

Finally, with dose equals to 1 the chart shows a clear difference and suggest
OJ may be related to a bigger teeth growth than VC.

## Hypothesis Test
### Dose effect on Length by Supplement Type
The T-Test shows that there is not a significant statistical difference when
comparing the effect that the supplement type had over tooth length:
```{r T-Test All Doses}
 test <- t.test( len~supp, data=ToothGrowth )
test
```

Confidence interval includes 0 which means it is possible that it is equal. Also
the p-value is `r round(test$p.value*100, 1)`% which is greater than 5%. The null 
hypothesis cannot be rejected.

### Supplement Type effect on Length by Dose
#### Dose Equals to 0.5
```{r T-Test Dose0.5}
data <- subset( ToothGrowth, dose %in% c(0.5) )
test <- t.test( len~supp, data=data )
```
P-Value of the test is `r round(test$p.value, 3)` which is lower than 5%. for
a dose equals to 0.5mm there is a significant statistic difference.


#### Dose Equals to 1
There seems to be an interesting relationship between the supplement type when
the dose is one.
```{r T-Test Dose1}
data <- subset( ToothGrowth, dose %in% c(1) )
test <- t.test( len~supp, data=data )
```
P-Value of the test is `r round(test$p.value, 3)` this is lower than 5%, too. 
There is also a significant statistic difference when the dose equals to 1mm .

#### Dose Between 0.5 and 1
```{r T-Test Dose 0.5 and 1}
data <- subset( ToothGrowth, dose %in% c(0.5, 1) )
t.test( len~supp, data=data )
```
Since p-value of the test is lower than 5% the t-test supports the alternative 
hypothesis. The effects of VC and OJ supplement types are different.


### Conclusion
There is a significant statistical difference in the Guinea Pig teeth growth between
VC and OJ supplement types when the dose is 0.5mg or 1mg. The second
plot clearly shows that OJ supplement type has better results than VC supplement
type.